---
title: "Brazilian Vaccine Data"
author: "Paul Truche"
date: "11/16/2020"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#tinytex::install_tinytex()
options(scipen = 999)
```

```{r, include=FALSE, warning=FALSE}
library(readxl)
imunizacoes_2019 <- read_excel("~/Desktop/imunizacoes 2019.xlsx")
imunizacoes_2019$Date<-as.Date(imunizacoes_2019$Date,
                        format = "%y/%m/%d")
```

```{r, include=FALSE, warning=FALSE}
library(qicharts)
library(wpp2019)
library(hrbrthemes)
library(ggplot2)
library(ggthemes)
library(data.table)
library(tidyverse)
library(scales)
library(plyr)
library(reshape)
```

In this study, we examine how the number of vaccines adminstered in Brazil has changed as a result of the COVID19 pandemic.
<br><br><br><br>

Our data comes from DATASUS which tabulates the number of vaccines distributed by state for each month.  Here we utilize data from 2015 thorugh 2020.
<br><br><br><br>

```{r, echo=FALSE, warning=FALSE,fig.cap="Total vacccines adminstered across Brazil"}
df<-data.frame(imunizacoes_2019)

# Most basic bubble plot
ggplot(df, aes(x=Date, y=Total.vaccines)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month") +
  ggtitle("Total Vaccines 2015:2020") +
  theme_fivethirtyeight()+
   scale_y_continuous(limits = c(3000000, 18000000)) +
  scale_color_fivethirtyeight()+
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-02-11")))
```
<br><br><br><br>

```{r, echo=FALSE, warning=FALSE}
dataforlong<-df[,c(1,3:7)]

dataforlong<-dataforlong %>% 
  dplyr::rename( North = North.vaccines, Northeast = Northeast.vaccines, Southeast =Southeast.vaccines , South =South.vaccines, Midwest =Midwest.vaccines)

dflong <- reshape::melt(dataforlong, id.vars="Date")

dflong<-dflong %>% 
  dplyr::rename(
    State = variable,
     Month = Date)

ggplot(dflong, aes(x=Month, y=value, group=State, color=State)) +
  geom_line( color="steelblue") + 
  geom_point() +
  xlab("Month") +
  ggtitle("Total By State 2015:2020") +
  theme_fivethirtyeight()+
  # scale_y_continuous(limits = c(0, 3000000)) +
  theme(axis.text.x=element_text(angle=60, hjust=1)) +
  scale_x_date(labels = date_format("%b-%Y"), date_breaks = "3 month",limit=c(as.Date("2015-01-01"),as.Date("2020-02-11")))
```
<br><br><br><br>

Here we use an I chart followed by a moving range (MR) chart

```{r, echo=FALSE, warning=FALSE}
df$Total2<-df$Total.vaccines/(df$Total.pop/1000)
# Create vector of random values to plot

nationaltotal<-qic(Total2,
    x     = Date,
    data  = df,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 population)',
    xlab  = '',
    nint=30,
     x.format = "%b-%Y", main  = "Total Vaccinations Brazil: 2015-2020",sub ="")

nationaltotal<-qic(Total2,
    x     = Date,
    data  = df,
    chart = 'i',
    ylab  = 'Distributed Vaccines (per 1000 population)',
    xlab  = '',
    nint=30,
     x.format = "%b-%Y", main  = "Total Vaccinations Brazil: 2015-2020",sub ="Break at March 2020", breaks=63)

qic(df$Total2,
    chart = 'mr',
    main  = 'Pairwise differences in Vaccine Rates (MR chart)',
    ylab  = 'Vaccines (Per 1000 pop)',
    xlab  = 'Month')
```
<br><br><br><br>
Below we use an xbar chart in order to incorperate vaccine rates per 1000 people by state.  
<br><br><br><br>

```{r, echo=FALSE, warning=FALSE}

df$Northadj<-df$North.vaccines/(df$North.pop/1000)
df$Northeastadj<-df$Northeast.vaccines/(df$Northeast.pop/1000)
df$Southeastadj<-df$Southeast.vaccines/(df$Southeast.pop/1000)
df$Southadj<-df$South.vaccines/(df$South.pop/1000)
df$Midwestadj<-df$Midwest.vaccines/(df$Midwest.pop/1000)

dataforlong2<-df[,c(1,15:19)]

dataforlong2<-dataforlong2 %>% 
  dplyr::rename( North = Northadj, Northeast = Northeastadj, Southeast =Southeastadj , South =Southadj, Midwest = Midwestadj)

dataforlong2 <- reshape::melt(dataforlong2, id.vars="Date")

dataforlong2<-dataforlong2 %>% 
  dplyr::rename(
    State = variable,
     Month = Date)

qic(value, 
    x     = Month, 
    data  = dataforlong2,
    chart = "xbar",
    main  = 'Average vaccination rate (Xbar chart; across states)',
    ylab  = 'Vaccines rate (per 1000 people)',
    xlab  = 'Month',x.format = "%b-%Y")

qic(value, 
    x     = Month, 
    data  = dataforlong2,
    chart = 'xbar',
    main  = 'Average vaccination rate (Xbar chart; across states)',
    ylab  = 'Vaccines rate (per 1000 people)',
     sub = "break at March 2020",
    xlab  = 'Month',breaks = 63,x.format = "%b-%Y")

```


